Abstract Spatial Specification of Stencil Computations

### 2016-06-23 paper.tex
explanation on the model

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\end{array}\right]}
\newcommand{\vthree}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\\$#3$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

@@

We give a model of our specifications which explains
the range of indices which correspond to the specifications.
The model is based on multi-sets since array indices my be repeated
in an expression.

Figure~\ref{fig:model} describes the model in terms of
an interpretation function $\interp{-}$ which maps specifications
to multisets of relative indices.

\begin{definition}Relative indices are drawn
from $(\mathbb{Z}_{\bot})^{\mathbb{D}}$, essentially an infinite
tuple of offsets. For example, $(i, j+1)$ corresponds to
the relative index $(0, 1, \bot, \bot, \ldots)$
\end{definition}

The main interpretation function is overloaded on lists of
variable-spec pairs $\interp{\overline{v : S}}$,
returning multisets of variable-relative-index pairs. This provides
the top-level definition of the model, with $\interp{-}$ overloaded
on $S$, $\interp{S}$ mqpping to multisets of
relative indices not associated to an array avariable.

@@

>>>>

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}
\newcommand{\vtwoh}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{cc}#1 \; & \; #2\end{array}\right]}
\newcommand{\vthreeh}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{ccc}#1 \; & \; #2 \; & \; #3\end{array}\right]}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\end{array}\right]}
\newcommand{\vthree}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\\$#3$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

@@

We provide a semantic model of the stencil specification language as
the basis for building correct implementations, showing the soundness
of the equational theory and approximations, and as a basis for the
inference, checking, and program synthesis algorithms.
The semantics of stencil specifications can be given a denotational
model as multisets of vectors, represent array indexing.

\begin{definition}
  A specification of maximum dimensionality $n$, is modelled by a
  multi-set of column vectors of size $n$  ($n$-dimensional vectors /
  $n$-vectors for short)
  with values drawn from $\mathbb{Z} \cup \{\infty\}$, representing
  relative indices. The $\infty$ element represents non-relative
  indices (\eg{}, constants), which are seen as infinite with respect
  to the frame of reference of a relative offset from an induction
  variable.

  For example, the array index $(i, j+1)$ corresponds to the vector
  $\vtwoh{0}{1}$, whereas, the $(0, j)$ corresponds to the
  vector $\vtwoh{\infty}{0}$ since $0$ is a non-relative index. 
\end{definition}

Figure~\ref{fig:model} outlines the core rules for regions
via the interpretation function $\interp{-}$.  We explain the
intermediate notations and definitions:

\begin{definition}A \emph{single-entry vector} of size $n$, denoted 
$\singleEntry{r}{n}$, is a vector 
where the $r^{th}$ entry is $1$ and all others are $0$, \eg{},
$\singleEntry{1}{2} = \vtwoh{1}{0}$.
\end{definition}

\begin{definition}The product of two models $\otimes$ is defined:
%%
\begin{align*}
M \otimes N & =
  \{(u \bowtie v)_i \mid i \in \{0, \ldots, 2^n \}, (u, v) \in
              M \times N\}^1
\end{align*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors).  Using this, the $\otimes$ operation constructs the set
of every pairwise permutation for every pair $(u, v)$ of $n$-vectors in the
Cartesian product of $M \times N$. Note the $^1$ superscript here 
on the set-builder notation which denotes that this construction
constructs a degenerate multiset (where all the multiplicities are 1).
This is important for the function of $\otimes$ in the model (see below).

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by 
non-determinstically picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:  
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{ccc}
0 & 1 & 2 \\
0 & 1 & 5 \\
0 & 4 & 2 \\
0 & 4 & 5 \\
3 & 1 & 2 \\
3 & 1 & 5 \\
3 & 4 & 2 \\
3 & 4 & 5 
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors 
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise pemutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_i + \neg b_{i,j} v_i
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer $i$. 
Recall that the vectors are defined over the domain
$\mathbb{Z} \cup \{\infty\}$. 
For addition and multiplication, we take 
 $\infty$ to be an absorbing element. Thus, if either
vector in a pairwise permutation contains $\infty$ in
the $j^{th}$ entry, then $j^{th}$ entry of every pairwise
permutation is $\infty$, \eg{}, 
%%
\begin{equation*}
\vtwoh{0}{1} \bowtie \vtwoh{2}{\infty} = 
\setlength{\arraycolsep}{0.5em}
\left[\begin{array}{cc}
0 & \infty \\
0 & \infty \\ 
2 & \infty \\
2 & \infty
\end{array}
\right]
\end{equation*}

%%
\end{definition}

%The main interpretation function is overloaded on lists of
%variable-spec pairs $\interp{\overline{v : S}}$,
%returning multisets of variable-relative-index pairs. This provides
%the top-level definition of the model, with $\interp{-}$ overloaded
%on $S$, $\interp{S}$ mqpping to multisets of
%relative indices not associated to an array avariable.

@@

\paragraph{Step 4: Convert spans into region AST}

Given the set $\textsf{regions}(M)$ of spans covering the indexing
space, the next step converts these into abstract syntax tree
for the specification language.
